---
title: "map read_excel reprex"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", error = TRUE)
```


# load packages
```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(here)
```

I have mutliple excel files that live in a data folder. I want to read them all into R, adding a filename variable to each. I found the original solution to the problem at this [stackflow link](https://stackoverflow.com/questions/47540471/load-multiple-excel-files-and-name-object-after-a-file-name)

# Original solution

```{r}
# define a path where the data lives

file_path <- here("data")

# get a list of files at that file_path
file_list <- list.files(path = file_path, pattern='*.xlsx')

# use map to read_excel all of those files, adding a file_name variable to each one

list_raw <- file_list %>% 
  map(~ read_excel(path = .x, range = "A1:C5")  %>%
        mutate("file_name" = .x))

```
This error had me stumped so I reproduced the problem using some dummy fruit data and made a git repo. Jenny Bryan VERY kindly came up with lots of different options for solving the problem. I know MUCH more about working with file paths as a result. 
I've added my two cents and Q and A to Jenny's solutions below in case these options are useful to anyone else. Thanks again, Jenny- you are the best!


# Trouble shooting the error 
## 6 options to deal with paths (by Jenny Bryan)

Here's a troubleshooting chunk that reveals the working directory and `file_list`. The path problem is going to be that `list.files()`, by default, only returns file names. But you need a path relative to whatever the working directory will be at render time or, even better, a resilient path determined at render time by here.

```{r}
getwd()

file_list

file.exists(file_list)
```


#### Option 1

Be explicit about the path, using `path = here("data", .x)` inside `read_excel()`

```{r}
file_list %>% 
  map(~ read_excel(path = here("data", .x), range = "A1:C5")  %>%
        mutate("file_name" = .x))

```

#### Option 2

Use `here("data", .x)` on `file_list` on the way in. Now you'll want to use `basename()` on the paths before adding to the data frame.

```{r}
file_list %>% 
  here("data", .) %>% 
  map(~ read_excel(path = .x, range = "A1:C5")  %>%
        mutate("file_name" = basename(.x)))

```

> Q: What does [basename](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/basename) do? 

> A: removes all of the path up to and including the last path separator (if any). Without it, the filename column ends up having the whole path listed (/Users/jenny/Desktop/map_excel_reprex/data/fruitdata1.xlsx)

#### Option 3 

Use `list.files(full.names = TRUE)` when capturing the file paths. Again, use `basename()` on the paths before adding to the data frame.

Remember- here is how I made the first file_list. It is just a list of file names. 
```{r}
file_list1 <- list.files(path = file_path, pattern='*.xlsx')

file_list1

```
Here is how JennyB suggests you make a file list. Specifying full.names = TRUE gets the whole path. 
```{r}

file_list2 <- list.files(path = here("data"), pattern='*.xlsx', full.names = TRUE)

file_list2
```

Using the file_list2 (with the whole path) and then basename (to strip the bit of the path you dont need in the filename column). 
```{r}

file_list2 %>% 
  map(~ read_excel(path = .x, range = "A1:C5")  %>%
        mutate("file_name" = basename(.x)))

```

#### Option 4

The `fs` package has useful functions for dealing with path issues. 

> Q: What does `fs` stand for?

> A: File systems; the `fs` package is a cross-platform interface to file system operations

Use `fs::dir_ls()` instead of `list.files()`, which IMO has better defaults than `list.files()`. Now we'll switch to `fs::path_file()` to shorten the paths before adding to the data frame.

```{r}
library(fs)

file_list_fs <- dir_ls(here("data"), glob = "*.xlsx")

file_list_fs
```
> Q: what is "glob"?

> A: glob patterns specify sets of filenames with wildcard characters, see [Wikipedia](https://en.wikipedia.org/wiki/Glob_(programming)

```{r}

file_list_fs %>% 
  map(~ read_excel(path = .x, range = "A1:C5")  %>%
        mutate("file_name" = fs::path_file(.x)))

```


#### Option 5

If `data` held *only* Excel files you could use `fs::dir_map()`. But it won't work as long as that `.Rmd` is in there.

```{r eval = FALSE}

foo <- function(full_path) {
  read_excel(path = full_path, range = "A1:C5") %>%
    mutate("file_name" = path_file(full_path))
}

dir_map("data", foo)
```


#### Option 6 

I suspect you will eventually want to row bind these data frames? In that case, I think the most pleasing work flow is with `purrr::map_dfr()`. If your paths are well-named from the start, its `id` argument accomplishes what you doing with `mutate()`.

```{r}
# define list of files including path
file_list_fs <- dir_ls(here("data"), glob = "*.xlsx")

# pull just the filenames from that list of paths using path_file()
names(file_list_fs) <- path_file(file_list_fs)

file_list_fs %>% 
  map_dfr(~ read_excel(path = .x, range = "A1:C5"), .id = "file_name")
```

> Q: I know that map_df() returns a dataframe, what does map_dfr() do?

> A: Requires dplyr to be installed and creates a df by row binding. There is also a map_dfc() function that does the same binding by columns. [More info here](https://purrr.tidyverse.org/reference/map.html)